---
title: "06_diet_classification"
output: html_document
date: "2025-02-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# packages 
library(sf)
library(dplyr)

```

## First we'll assign some latrine metrics/information 
- Classify by freshwater/saltwater 
- Calculate distance to saltwater 
- Assign water body names
- Assign latrine types 
- Combine above with spatial covariates 
```{r}
####################################

# let's classify latrines by distance to freshwater 

### aquatic 
# read in water
water <- st_read(here("spatial_data", 
                      "DNR_Hydrography_-_Water_Bodies_-_Forest_Practices_Regulation.shp")) 

sf_use_s2(FALSE)
st_crs(water)

# crop to our map area and put in the projection we're working in
water <-  water %>%
  st_transform(crs = 4326) %>% 
  st_crop(c(xmin= -121.7, ymin = 46.7, xmax = -122.8, ymax = 48.5)) %>%
  st_transform(crs = 32610) 

# read in our site locations 
sites <- st_read(here("clean_data", "latrines.shp")) %>% 
  st_transform(crs = "EPSG:32610")


# let's add a new attribute to water that will be based on hydrography categories 
# according to the metadata https://fortress.wa.gov/dnr/adminsa/GisData/metadata/ROPA_WBHYDRO.pdf
# and the shoerline data set https://geo.wa.gov/datasets/wadnr::shorezone-inventory-salt-marsh/explore?location=47.903540%2C-122.176510%2C10.00
# all saltwater bodies are encompassed under "bay"

water <- water %>%
  mutate(water_type = ifelse(WB_HYDR_FT == "Bay", "saltwater", "freshwater"))


# now let's calculate distance to saltwater for each site
# this is a straight line closest distance 

saltwater <- water %>% filter(water_type == "saltwater")

# calculate the nearest distance from each site to the closest saltwater feature
sites <- sites %>%
  mutate(saltwater_dist = st_distance(geometry, st_geometry(saltwater)) %>%
           apply(1, min))  # get the minimum distance for each point

print(sites)




# now let's concatenate sites with our other spatial covariates 

all_covs <- read_csv(here("spatial_data", "covariates", "ALL_ENV_URB_SITES_2000m.csv"))


site_covs <- left_join(sites, all_covs, by = c("city", "site")) %>%
  select(-c(1:5,8:13))

colnames(site_covs)

# write_csv(site_covs, here("spatial_data", "covariates", "ALL_ENV_URB_WA_SITES_2000m.csv"))

```


## Now we'll create some diet categories/classifications

First read in diet data.
```{r}

verts <- read_csv(here("clean_data", "12S_final.csv"))
inverts <- read_csv(here("clean_data", "16S_final.csv"))
site_covs <- read_csv(here("spatial_data", "covariates", "ALL_ENV_URB_WA_SITES_2000m.csv"))

```


Next we'll classify diet items into functional groups for GLMM analysis. 
```{r pressure, echo=FALSE}

#################
# diet classifications 
# for freshwater/saltwater, checked using https://www.sealifebase.se/
## freshwater malacostracae
# Asellidae (isopod), Cambaridae (crayfish), Astacidae (crayfish)

fresh_mala_fam <- c("Asellidae", "Cambaridae", "Astacidae")

inverts <- inverts %>%
  mutate(diet_class = ifelse(family %in% fresh_mala_fam, "fresh_mala", NA))

sum(inverts$diet_class == "fresh_mala", na.rm = TRUE)


## saltwater malacostracae
# MANUAL LIST: 
# Amphiphoidae (amphipods), Aoridae (amphipod), Cancridae (crabs), Caprellidae (amphipod), 
# Corophiidae (amphipod -- salt and fresh), Dulichiidae (amphipod), Ischyroceridae (amphipod), 
# Sphaeromatidae (isopod), Varunidae (crabs), Panopeidae (mud crabs), Oregoniidae (crabs), 
# Paguridae (hermit crabs), Pandalidae (shrimps), Cheiragonidae (helmet crab)


inverts %>%
  filter(class == "Malacostraca") %>%
  pull(family) %>%
  unique()

salt_mala_fam <- c("Amphiphoidae", "Aoridae", "Cancridae", "Caprellidae", 
                   "Corophiidae", "Dulichiidae", "Ischyroceridae", "Sphaeromatidae", 
                   "Varunidae", "Panopeidae", "Oregoniidae", "Paguridae", 
                   "Pandalidae", "Cheiragonidae")

inverts <- inverts %>%
  mutate(diet_class = ifelse(family %in% salt_mala_fam, "salt_mala", diet_class))

sum(inverts$diet_class == "salt_mala", na.rm = TRUE)

## insects
insects <- c("Insecta") 

inverts <- inverts %>%
  mutate(diet_class = ifelse(class %in% insects, "insects", diet_class))

sum(inverts$diet_class == "insects", na.rm = TRUE)

## gastropods
gastropods <- c("Gastropoda") 

inverts <- inverts %>%
  mutate(diet_class = ifelse(class %in% gastropods, "gastropods", diet_class))

sum(inverts$diet_class == "gastropods", na.rm = TRUE)

#########
# vert classifications 

## ray finned fishes 

# centrarchiformes, perciformes, blenniiformes (perch-like + surf perch)
# cypriniformes (cyprinid fish -- carps and minnows)
# siluriformes (catfish -- all brown bullhead)
# clupeiformes (pacific herring -- not enough to isolate)
# pleuronectiformes (flatfish -- 2 spp)
# salmoniformes 

verts$class

fish <- c("Actinopteri") 

verts <- verts %>%
  mutate(diet_class = ifelse(class %in% fish, "fish", NA))

sum(verts$diet_class == "fish", na.rm = TRUE)

verts %>%
  filter(class == "Actinopteri") %>%
  pull(order) %>%
  unique()


# breaking down groups of ray-finned fish

# perch-like 
perch <- c("Blenniiformes", "Perciformes", "Cantrarchiformes") 

verts <- verts %>%
  mutate(diet_class = ifelse(order %in% perch, "perch", diet_class))

sum(verts$diet_class == "perch", na.rm = TRUE)

# cypriniformes
cypri <- c("Cypriniformes") 

verts <- verts %>%
  mutate(diet_class = ifelse(order %in% cypri, "cypri", diet_class))

sum(verts$diet_class == "cypri", na.rm = TRUE)

# siluriformes (brownbullhead)
bullhead <- c("Siluriformes") 

verts <- verts %>%
  mutate(diet_class = ifelse(order %in% bullhead, "bullhead", diet_class))

sum(verts$diet_class == "bullhead", na.rm = TRUE)

# pleuronectiformes
flatfish <- c("Pleuronectiformes") 

verts <- verts %>%
  mutate(diet_class = ifelse(order %in% flatfish, "flatfish", diet_class))

sum(verts$diet_class == "flatfish", na.rm = TRUE)

unique(verts$diet_class)

## ducks  & rails 
verts %>%
  filter(class == "Aves") %>%
  pull(family) %>%
  unique()

ducksrails <- c("Anatidae", "Rallidae") 

verts <- verts %>%
  mutate(diet_class = ifelse(family %in% ducksrails, "ducksrails", diet_class))

sum(verts$diet_class == "ducksrails", na.rm = TRUE)

## passerine birds 

passerine <- c("Corvidae", "Turdidae") 

verts <- verts %>%
  mutate(diet_class = ifelse(family %in% passerine, "passerine", diet_class))

sum(verts$diet_class == "passerine", na.rm = TRUE)

## small mammals -- non-anthro
verts %>%
  filter(class == "Mammalia") %>%
  pull(family) %>%
  unique()

smallmammals <- c("Muridae", "Castoridae", "Leporidae", "Cricetidae")

verts <- verts %>%
  mutate(diet_class = ifelse(family %in% smallmammals, "smallmammals", diet_class))

sum(verts$diet_class == "smallmammals", na.rm = TRUE)

## anthropogenic mammals 
# turkey, chicken, sheep, cow, cat, pig, muscovy duck 
anthro <- c("Bos" , "Phasianidae" , "Oves" , "Gallus", "Felis", "Sus", "Cairina")

verts <- verts %>%
  mutate(diet_class = ifelse(genus %in% anthro, "anthro", diet_class))

sum(verts$diet_class == "anthro", na.rm = TRUE)

## salmonids (overwrite salmonids in ray finned)

salmon <- c("Salmonidae") 

verts <- verts %>%
  mutate(diet_class = ifelse(family %in% salmon, "salmonids", diet_class))

sum(verts$diet_class == "salmonids", na.rm = TRUE)


############## 
## combine all data 
inverts$processed_taxon
inverts <- inverts %>% 
  select(-possible_taxon,superclass)

# Get the full set of column names from both data frames
all_columns <- union(colnames(verts), colnames(inverts))

# Ensure verts has all columns by adding missing ones with NA
for (col in setdiff(all_columns, colnames(verts))) {
  verts[[col]] <- NA
}

# Ensure inverts has all columns by adding missing ones with NA
for (col in setdiff(all_columns, colnames(inverts))) {
  inverts[[col]] <- NA
}

# Reorder columns to match the full set
verts <- verts %>% select(all_columns)
inverts <- inverts %>% select(all_columns)

# Bind the two data frames together
all_diet <- as.data.frame(bind_rows(verts, inverts))
class(all_diet)

```

Next we'll classify all diet categories into one of four categories for visualization purposes, as well as classifying species into native or non-native. 

During this step we'll also throw out some diet items that we don't think are ecologically relevant. 

We'll create a new data set, as many taxa will be dropped. 

The four categories will be: saltwater aquatic organisms, freshwater aquatic organisms, waterfowl, mammals. We'll classify each scat going class to class. 

We'll start with invertebrates, see comments below for reasoning on which to drop: 
```{r}
head(all_diet)

## insecta

# list all distinct families in Insecta 
all_diet %>% filter(class == "Insecta") %>% 
  distinct(family) # we only have to go down to family for insects

# hydrophilidae -- freshwater beetles

# freshwater insects to keep 
fresh_insects <- c(
  "Hydrophilidae",  # freshwater beetles
  "Carabidae", # freshwater beetles 
  "Dystiscidae", # freshwater beetles 
  "Staphylinidae", # freshwater rove beetles 
  'Tachinidae', # parasitoid freshwater flies 
  "Coenagrionidae", # damselflies 
  "Micronectidae" # water boatman
)

# terrestrial arthropods  to keep
terr_insects <- c(
  "Diplopoda", # millipedes 
  "Gryllidae", # crickets 
  "Curculionidae" # weevils (poplar & willow borer) -- NON NATIVE 
)

# insects to toss for now -- although they may be eating larvae of this species 
toss_insects <- c("Chironomidae", # midges -- larvae in winter/spring
                   "Culicidae", # mosquitoes -- larvae all summer 
                   "Psychodidae", # moth flies 
                   "Therevidae", # stiletto flies -- larvae in spring/summer 
                   "Coccinellidae", # ladybugs, only one occurrence 
                   "Panorpodidae", # scorpion flies, only one occurrence 
                   NA# nothing 
)

# create a new data set
new_diet <- all_diet %>% 
  mutate(freshwater = ifelse(family %in% fresh_insects, 1, 0)) # fill all rest with 0s -- only for this first one

new_diet <- new_diet %>% 
  mutate(terrestrial = ifelse(family %in% terr_insects, 1, 0)) # fill all rest with 0s -- only for this first one

## crustaceans 
# we already did a salt/freshwater breakdown for malacostracans so we'll use that 
all_diet %>% filter(subphylum == "Crustacea") %>% 
  distinct(class)

# ignore all other classes for now 

new_diet <- new_diet %>% 
  mutate(freshwater = ifelse(family %in% fresh_mala_fam , 1, freshwater))

new_diet <- new_diet %>% 
  mutate(saltwater = ifelse(family %in% salt_mala_fam , 1, 0)) # first one so we'll fill with 0s


# platyhelminths 
# ligula intestinalis will go under saltwater 

new_diet <- new_diet %>% 
  mutate(saltwater = ifelse(phylum %in% "Platyhelminthes", 1, saltwater)) 

# mollusks 
all_diet %>% filter(phylum == "Mollusca") %>% 
  distinct(class)

all_diet %>% filter(class == "Gastropoda") %>% 
  distinct(order)

# we'll have to drop all gastropods since we don't have any data on them past class 

all_diet %>% filter(class == "Bivalvia") %>% 
  distinct(order)

all_diet %>% filter(order == "Venerida") %>% 
  distinct(species)

fresh_bivalve <-  c(
"Venerida" # all fingernail clam 
)

new_diet <- new_diet %>% 
  mutate(freshwater = ifelse(family %in% fresh_bivalve, 1, freshwater)) 


## remaining invertebrates 
# we're going to throw out all other invertebrates 
# branchiopods because they're all daphnia 
# hexanauplia & ostracoda bc they only had 1-2 and only amplified to family
# ones that have NA for family because we dont have enough info 
# arachnida because they're bird mites 
# collembola bc they're globular springtails

remove_inverts <- c("Branchiopoda", "Hexanauplia", "Ostracoda", "Arachnida", "Collembola", "Gastropoda")

new_diet <- new_diet %>%
  filter(
    !class %in% remove_inverts,  
    !is.na(class),                  
    !is.na(family)                  
  )

# let's also start a new column for native vs. non-native inverts, and add our non-native inverts to it 
non_native_insects_sp <- c("Aoroides columbiae",
                           "Procambarus clarkii",
                           "Faxonius virilis",
                           "Oronectes rusticus",
                           "Acrosternum hilare",
                           "Rhagonycha fulva",
                           "Cipangopaludina chinensis"
                        )

non_native_insects_gen <- c("Cryptorynchus",
                            "Tetramorium"
)

new_diet <- new_diet %>% 
  mutate(non_native = ifelse(species %in% non_native_insects_sp, 1, 0))    # 0 for first one

new_diet <- new_diet %>% 
  mutate(non_native = ifelse(genus %in% non_native_insects_gen, 1, non_native))    
  
```

Now vertebrates, starting with fish -- in this data set all fish were in class Actinopteri. 

```{r}

list(new_diet %>% filter(class == "Actinopteri") %>% 
  distinct(family))

# family cottidae
list(new_diet %>% filter(family == "Cottidae") %>% 
  distinct(species))

list(new_diet %>% filter(family == "Cottidae") %>% 
  distinct(genus))


fresh_fish_gen <- c("Cottus") # freshwater sculpins

salt_fish_sp <- c("Scorpaenichthys marmoratus",  # cabezon
               "Leptocottus armatus", # pacific staghorn sculpin
               "Clinocottus acuticeps", # sharpnose sculpin
               "Hemilepidotus hemilepidotus") # red irish lord

# family psychrolutidae
list(new_diet %>% filter(family == "Psychrolutidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family == "Psychrolutidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Dasycottus setiger") # spinyhead sculpin

# family gasterosteidae
list(new_diet %>% filter(family == "Gasterosteidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family == "Gasterosteidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Gasterosteus aculeatus") # three-spined stickleback, BOTH fresh and salt 
fresh_fish_sp <- c(fresh_fish_sp, "Gasterosteus aculeatus") # three-spined stickleback, BOTH fresh and salt 

# family centrarchidae
list(new_diet %>% filter(family == "Centrarchidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family == "Centrarchidae") %>% 
  distinct(species))

fresh_fish_sp <- c(fresh_fish_sp, "Ambloplites rupestris",# rock bass -- NON-NATIVE 
                   "Lepomis cyanellus", # green sunfish -- NON-NATIVE 
                   "Lepomis gibbosus", # pumpkinseed sunfish-- NON-NATIVE, considered invasive, maybe brackish too ?
                  "Micropterus dolomieu", # smallmouth bass -- NON-NATIVE
                  "Micropterus salmoides", # largemouth bass -- NON-NATIVE 
                  "Pomoxis nigromaculatus" # black crappie -- NON-NATIVE 
                  ) 

non_native_fish <- c("Ambloplites rupestris", "Lepomis cyanellus","Lepomis gibbosus","Micropterus dolomieu","Micropterus salmoides","Pomoxis nigromaculatus")

# family ictaluridae
list(new_diet %>% filter(family == "Ictaluridae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Ictaluridae") %>% 
  distinct(species))

fresh_fish_sp <- c(fresh_fish_sp, "Ameiurus nebulosus")  # brown bullhead catfish -- NON-NATIVE 

non_native_fish <- c(non_native_fish, "Ameiurus nebulosus")

# family cyprinidae 
list(new_diet %>% filter(family == "Cyprinidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Cyprinidae") %>% 
  distinct(species))

fresh_fish_sp <- c(fresh_fish_sp, "Cyprinus carpio")  # common carp -- NON-NATIVE 

non_native_fish <- c(non_native_fish, "Cyprinus carpio")

# family cobitidae
list(new_diet %>% filter(family == "Cobitidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Cobitidae") %>% 
  distinct(species))

fresh_fish_sp <- c(fresh_fish_sp, "Misgurnus anguillicaudatus")  # pond loach -- NON-NATIVE 

non_native_fish <- c(non_native_fish, "Misgurnus anguillicaudatus")

# family leuciscidae
list(new_diet %>% filter(family == "Leuciscidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Leuciscidae") %>% 
  distinct(species))

fresh_fish_sp <- c(fresh_fish_sp, "Mylocheilus caurinus", # peamouth
                   "Rhinichthys cataractae" #longnose dace minnow
                   )

# family percidae
list(new_diet %>% filter(family == "Percidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Percidae") %>% 
  distinct(species))

fresh_fish_sp <- c(fresh_fish_sp, "Perca flavescens" # yellow perch -- NON-NATIVE 
                   )

non_native_fish <- c(non_native_fish, "Perca flavescens")

# family catostomidae
list(new_diet %>% filter(family == "Catostomidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Catostomidae") %>% 
  distinct(species))

fresh_fish_sp <- c(fresh_fish_sp, "Catostomus macrocheilus" # largescale sucker 
                   )

# family stichaeidae
list(new_diet %>% filter(family == "Stichaeidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Stichaeidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Anoplarchus insignis", # slender cockscomb 
                  "Lumpenus sagitta",  # snake prickleback
                  "Xiphister atropurpureus" # black prickleback
                  )

# family Salmonidae
list(new_diet %>% filter(family == "Salmonidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Salmonidae") %>% 
  distinct(species))

# salmon are anadromous so they will be BOTH fresh and salt 
salt_fish_sp <- c(salt_fish_sp, "Oncorhynchus tshawytscha", # chinook salmon
                  "Oncorhynchus clarkii", # cutthroat trout 
                  "Oncorhynchus gorbuscha", # pink salmon
                 "Oncorhynchus kisutch" # coho salmon
                  )

salt_fish_sp <- c(salt_fish_sp, "Oncorhynchus tshawytscha",
                  "Oncorhynchus clarkii",
                  "Oncorhynchus gorbuscha",
                  "Oncorhynchus kisutch"
                  )


# family Anarhichadidae
list(new_diet %>% filter(family == "Anarhichadidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Anarhichadidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Anarrhichthys ocellatus" # wolf-eel
                  )

# family Pleuronectidae
list(new_diet %>% filter(family == "Pleuronectidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Pleuronectidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Lepidopsetta bilineata", # pacific rock sole 
                  "Platichthys stellatus",# starry flounder
                  "Pleuronichthys coenosus" # C-O sole
                  )

# family Clupeidae
list(new_diet %>% filter(family == "Clupeidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Clupeidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Clupea pallasii" # pacific herring
                  )


# family Pholidae
list(new_diet %>% filter(family == "Pholidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Pholidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Pholis ornata", # saddleback gunnel
                  "Apodichthys flavidus" # penpoint gunnel
)
                  

# family Hexagrammidae
list(new_diet %>% filter(family == "Hexagrammidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Hexagrammidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Ophiodon elongatus", # lingcod
                  "Hexagrammos lagocephalus" # rock greenling
)
              

# family Jordaniidae
list(new_diet %>% filter(family == "Jordaniidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Jordaniidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Jordania zonope" # longfin sculpin
                  )

# family Agonidae
list(new_diet %>% filter(family == "Agonidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Agonidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Agonopsis vulsa", # northern spearnose poacher
                  "Blepsias cirrhosus" # silverspotted sculpin
                  )

# family Embiotocidae
list(new_diet %>% filter(family == "Embiotocidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Embiotocidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Cymatogaster aggregata" # shiner perch 
                  )

# family Cyclopsettidae
list(new_diet %>% filter(family == "Cyclopsettidae") %>% 
  distinct(genus))

list(new_diet %>% filter(family =="Cyclopsettidae") %>% 
  distinct(species))

salt_fish_sp <- c(salt_fish_sp, "Citharichthys stigmaeus" # speckled sanddab
                  )

```

Next we'll do the same with mammals.
```{r}

list(new_diet %>% filter(class == "Mammalia") %>% 
  distinct(family))

list(new_diet %>% filter(family == "Mustelidae") %>% 
  distinct(genus))
# almost all families are associated with terrestrial environments, except beavers (Castoridae) and mink (Mustelidae)
terr_mamm_fam <- c("Felidae", "Muridae") 

fresh_mamm_fam <- c("Mustelidae", "Castoridae")

```


Now birds. For now we'll classify all waterfowl that exists in both fresh & marine habitats in WA in both groups. 

```{r}
list(new_diet %>% filter(class == "Aves") %>% 
  distinct(family))


# ducks and rails 
salt_birds <- c("Anas acuta", # northern pintail -- fresh and marsh/brackish
                "Histrionicus histrionicus", # harlequin duck -- sea duck but also fresh 
                "Mareca americana", # american wigeon -- both 
                "Anas crecca", # green-winged teal -- both 
                "Clangula hyemalis", # long-tailed duck; sea duck, only in marine in WA 
                )

fresh_birds <- c("Anser albifrons", # greater white fronted goose -- freshwater, rare, passes through in spring migration ! 
                 "Cairina moschata" # muscovy duck -- NON-NATIVE domestic -- freshwater
                 "Fulica americana" # american coot -- mostly freshwater 
                 "Gruis canadensis" # sandhill crane -- mostly freshwater, may just be coots with seq error tbh
                 )

# passerines & chickens -- we'll classify as terrestrial 
terr_birds <- c("Gallus gallus",
                "Turdus migratorius",
                "Corvus brachyrhynchos")

# non native birds
non_native_birds <- c("Cairina moschata") 


```

Now amphibians.
```{r}
list(new_diet %>% filter(class == "Amphibia") %>% 
  distinct(family))


fresh_amph <- c("Aquarana catesbeiana", # american bullfrog -- NON-NATIVE 
                "Ambystoma"
)

non_native_amph <- c("Aquarana catesbeiana")

              


```

Now Hyperoartia (lampreys).
```{r}

```

Now assign vertebrate habitat classifications to the data set.
```{r}

new_diet <- new_diet %>% 
  mutate(freshwater = ifelse(family %in% fresh_fish_sp, 1, freshwater)) %>% 
  mutate(freshwater = ifelse(family %in% fresh_fish_gen, 1, freshwater))  %>% 
  mutate(freshwater = ifelse(family %in% fresh_mamm_fam, 1, freshwater)) 

new_diet <- new_diet %>% 
  mutate(saltwater = ifelse(family %in% salt_fish_sp, 1, saltwater)) 

new_diet <- new_diet %>% 
  mutate(terrestrial = ifelse(family %in% terr_mamm_fam, 1, terrestrial)) 


non_native_fish
non_native_insects_sp
non_native_insects_gen
non_native_amph

```


Remove all rows that don't have class or family assigned. 
```{r}
# now remove all rows where we don't have class or family

new_diet <- new_diet %>%
  filter(
    !is.na(class),                  
    !is.na(family)                  
  )

```

Now assign diet classifications. 
```{r}
new_diet <- new_diet %>%
  mutate(
    fish = ifelse(diet_class == "fish", 1, ifelse(is.na(diet_class), 0, 0)),
    ducksrails = ifelse(diet_class == "ducksrails", 1, ifelse(is.na(diet_class), 0, 0)),
    anthro = ifelse(diet_class == "anthro", 1, ifelse(is.na(diet_class), 0, 0)),
    smallmammals = ifelse(diet_class == "smallmammals", 1, ifelse(is.na(diet_class), 0, 0)),
    salmonids = ifelse(diet_class == "salmonids", 1, ifelse(is.na(diet_class), 0, 0)),
    passerine = ifelse(diet_class == "passerine", 1, ifelse(is.na(diet_class), 0, 0)),
    fresh_mala = ifelse(diet_class == "fresh_mala", 1, ifelse(is.na(diet_class), 0, 0)),
    salt_mala = ifelse(diet_class == "salt_mala", 1, ifelse(is.na(diet_class), 0, 0)),
    dungeness = ifelse(diet_class == "dungeness", 1, ifelse(is.na(diet_class), 0, 0)),
    perch = ifelse(diet_class == "perch", 1, ifelse(is.na(diet_class), 0, 0)),
    cypri = ifelse(diet_class == "cypri", 1, ifelse(is.na(diet_class), 0, 0)),
    bullhead = ifelse(diet_class == "bullhead", 1, ifelse(is.na(diet_class), 0, 0)),
    flatfish = ifelse(diet_class == "flatfish", 1, ifelse(is.na(diet_class), 0, 0)),
    insects = ifelse(diet_class == "insects", 1, ifelse(is.na(diet_class), 0, 0)),
    gastropods = ifelse(diet_class == "gastropods", 1, ifelse(is.na(diet_class), 0, 0)),
  ) %>% 
  mutate(across(starts_with("fish"):starts_with("gastropods"), ~ replace(., is.na(.), 0)))  


# let's save our final diet data set 
write_csv(new_diet, here("clean_data", "combined_diet_16s_12s_ms.csv"))
```

Finally, we'll bind the diet data back to the covariate data. 
```{r}

# let's add covariates to diet data 
# first we need to add site names to the diet list 

## re-read in all the data 

new_diet <- read_csv(here("clean_data", "combined_diet_16s_12s_ms.csv"))
site_covs <- read_csv(here("spatial_data", "covariates", "ALL_ENV_URB_WA_SITES_2000m.csv"))

metadata <- read_csv(here("data_from_ellie", "field_site_data.csv")) %>% 
  rename(sample_ID = "Scat ID",
         site = "Site",
         date = "Date", 
         month = "Month", 
         city = "City") %>% 
  select(sample_ID, site, date, month, city)

new_diet <- left_join(new_diet, metadata, by = c("sample_ID"))

# now join the spatial data
new_diet <- left_join(new_diet, site_covs, by = c("city", "site"))


# for some reason LCP047 is missing site data 
# it's from FGL so we'll fill the columns based on that 
colnames(new_diet)

new_diet <- new_diet %>%
  mutate(site = ifelse(grepl("LC047", sample_ID), "FGL", site)) %>%
  group_by(site) %>%
  mutate(across(35:62, ~ ifelse(sample_ID == "LC047" & is.na(.), first(na.omit(.)), .))) %>%
  ungroup()


# we'll change cornet bay to match the others 
"Cornet Bay" "CBW"

# we'll also add water body names to each site 
all_diet <- all_diet %>% 
  mutate(waterbody = case_when(
    site %in% c("LNM", "LWM", "MBB", "LSM", "LBP", "WNP", "CBM") ~ "Lake Washington",
    site %in% c("FGL", "DSS", "CWG", "BRF") ~ "Duwamish/Green River",
    site %in% c("BSW", "WPA") ~ "Portage & Union Bays",
    site %in% c("Cornet Bay", "GGP", "DNB") ~ "Puget Sound", 
    site %in% c("RVL", "GGC") ~ "Kelsey Creek")
  )

# we'll also add latrine type 
all_diet <- all_diet %>% 
  mutate(latrinetype = case_when(
    site %in% c("LNM", "LWM", "LSM", "CBM", "BSW", "WPA") ~ "Docks and marinas",
    site %in% c("FGL", "GGC") ~ "Golf courses",
    site %in% c("WNP", "DSS", "CWG", "BRF",  "CBW", "DNB", "MBB", "RVL", "GGP") ~ "Natural shoreline")
  )

# let's make our variables factors 
all_diet$waterbody <- as.factor(all_diet$waterbody)
all_diet$site <- as.factor(all_diet$site)

write_csv(all_diet, here("clean_data", "combined_diet_covs_all_ms.csv"))
```


